Ab Initio Effective One-Electron Potential Operators. II. Applications for Exchange-Repulsion Energy in Effective Fragment Potentials

### 2019-12-10 papers/3_oep_technique.Pauli/paper-r0.tex
Figures added to paper II

@@

Three complexes: 
(i) (H$_2$O)$_2$, 
(ii) H$_2$O--HOCH$_3$, 
(iii) (DMSO)$_2$,
were chosen as model systems to analyze the asymptotic dependence 
of EXR energy, which is shown in Figure~XXX. 
The reference (zero\hyp{}displacement) geometries
were obtained as described in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} 
and the structures  
along with the translation direction are depicted
in the insets in Figure~XXX.
Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
Small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO were developed
to minimize the computational cost of evaluating the summation terms
from Eq.~\eqref{e:rep.murrell-etal.S1.oep}. For this purpose, the basis set optimization 
as described in the Appendix~A in Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
in conjunction with the basin hopping global optimization algorithm,  %cite
available in the SciPy Python module. %cite

>>>>

The alternative EFP2 formulation %with OEP\hyp{}based first\hyp{}order term
is therefore still given by Eq.~\eqref{e:exr-efp2}
%\begin{equation} \label{e:exrep-new}
%E^{\rm Ex-Rep} \approx E^{\rm Ex} + E^{\rm Rep}_{\rm OEP} + E^{\rm Rep}(S^{-2}) \;,
but with $E^{\rm Rep}(S^{-1})$ replaced by OEP\hyp{}based
first\hyp{}order term from Eq.~\eqref{e:rep.murrell-etal.S1.oep}.
%with $E^{\rm Ex}$ given by Eq.~\eqref{e:exc-efp2}, 
%$E^{\rm Rep}(S^{-1})$ given by Eq.~\eqref{e:rep.murrell-etal.S1.oep} and
%$E^{\rm Rep}(S^{-2})$ given by Eq.~\eqref{e:rep-efp2.s2}.

@@

We implemented EXR energy models from Eq.~\eqref{e:exr-efp2}
(EFP2), Eq.~\eqref{e:rep.murrell-etal.S1.oep} (OEP), as well as the repulsion term of Murrell et~al.
from Eq.~\eqref{e:rep.murrell-etal.S1} (here abbreviated as S1-Ref) for validation of the EDF technique.
Benchmark EXR energy were estimated from the density decomposition scheme 
at the Hartree\hyp{}Fock level of theory (here referred as to the DDS/HF) 
developed by Mandado and Hermida\hyp{}Ram{\'o}n,\cite{Mandado.Hermida-Ramon.JCTC.2011} 
which is equivalent to the exact EXR energy 
in the intermolecular perturbation theory with exchange of Hayes and Stone
at this level of theory.\cite{Hayes.Stone.MolPhys.1984} 
To localize molecular orbitals, procedure that is necessary for the EFP2 calculations,
the Boys method\cite{Boys.RevModPhys.1960} was used. Note here that, since
Murrell et~al.'s theory is invariant with respect to unitary transformation
of molecular orbitals, Eq.~\eqref{e:rep.murrell-etal.S1.oep} is also invariant and does not require
orbital localization. 
%in which
%
%\begin{equation}\label{e:dds-exr}
% \Delta E^{\rm Ex-Rep}_{\rm DDS/HF} = 
%\end{equation}
%

Three complexes: 
(i) (H$_2$O)$_2$, 
(ii) H$_2$O--HOCH$_3$, 
(iii) (DMSO)$_2$,
were chosen as model systems to analyze the asymptotic dependence 
of EXR energy, which is shown in Figure~XXX. 
The reference (zero\hyp{}displacement) geometries
were obtained as described in Ref.\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} 
and the structures  
along with the translation direction are depicted
in the insets in Figure~XXX.
Energy\hyp{}optimizations were performed at the HF/6-31+G(d,p) level,
as implemented in 
the {\sc Gaussian16} quantum chemistry program package.\cite{Gaussian16}
Small auxiliary basis sets for H$_2$O, CH$_3$OH and DMSO were developed
to minimize the computational cost of evaluating the summation terms
from Eq.~\eqref{e:rep.murrell-etal.S1.oep}. For this purpose, the basis set optimization 
as described in Appendix~A to Ref.~\cite{Blasiak.Bednarska.Choluj.Bartkowiak.JCP.2019} was implemented
in conjunction with the basin hopping global optimization algorithm,  %cite
available in the SciPy Python module. %cite

{
\renewcommand{\arraystretch}{1.4}
\begin{table}[h]
\caption[]
{{\bf CPU timings in miliseconds of exchange\hyp{}repulsion single point 
energy calculations via various methods\footnotemark[1]}
}
\label{t:5}
\begin{ruledtabular}
\begin{tabular}{llccccccccc}
                      && \multicolumn{2}{c}{(H$_2$O)$_2$\footnotemark[2]} 
                      && \multicolumn{2}{c}{H$_2$O--HOCH$_3$\footnotemark[2]} 
                      && \multicolumn{2}{c}{(DMSO)$_2$\footnotemark[3]} \\
\cline{3-4}
\cline{6-7}
\cline{9-10}
Murrell et al.\footnotemark[4] && 3$\times 10^4$  & (10.0)  
                               && 3$\times 10^4$  & (10.0)
                               && 3$\times 10^4$  & (10.0) \\
Hayes-Stone\footnotemark[5]    && 3$\times 10^4$  & (10.0)  
                               && 3$\times 10^4$  & (10.0)
                               && 3$\times 10^4$  & (10.0) \\
EFP2                           && 3$\times 10^4$  & (10.0)  
                               && 3$\times 10^4$  & (10.0)
                               && 3$\times 10^4$  & (10.0) \\
OEP/large\footnotemark[6]      && 3$\times 10^4$  & (10.0)  
                               && 3$\times 10^4$  & (10.0)
                               && 3$\times 10^4$  & (10.0) \\
OEP/mini\footnotemark[6]       && 3$\times 10^4$  & (10.0)  
                               && 3$\times 10^4$  & (10.0)
                               && 3$\times 10^4$  & (10.0) \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{1.2 GHz AMD EPYC\texttrademark{} 7301 16-Core Processor, calculations performed on 1 core. 
Exchange\hyp{}repulsion energies are given in parentheses for reference (kcal/mol).
See also the implementation details in the main text.}
\footnotetext[2]{Primary basis set: 6-311++G(d,p).}
\footnotetext[3]{Primary basis set: 6-31++G(d,p).}
\footnotetext[4]{Reference~\cite{Murrell.Randic.Williams.Longuet-Higgins.ProcRSocLondA.1965}.}
\footnotetext[5]{Reference~\cite{Hayes.Stone.MolPhys.1984}.}
\footnotetext[6]{Eq.~\eqref{}. Auxiliary basis sets for the EDF in $E^{\rm Rep}(S^{-1})$ 
from Eq.~\eqref{e:rep.murrell-etal.S1.oep}
are: `large' - aug-cc-pVDZ-jkfit for (H$_2$O)$_2$ and H$_2$O--HOCH$_3$ systems as well as
aug-cc-pVDZ-ri for (DMSO)$_2$ system;
`mini' - optimized small basis sets (see Supporting Information).}
%
\end{table}
}
%
